---
title: 'LANDIS-II Biomass Succession pixel-level simulations: Southern Interior Mountains
  - Interior Cedar Hemlock (preliminary results)'
author: "Dominic cyr"
output:
  pdf_document: default
  html_document:
    keep_md: yes
    theme: cosmo
    toc: yes
---
Updated on `r format(Sys.time(), "%b %d %Y")`

-------

```{r packageLoadingChunk, include=FALSE}
require(RCurl)
require(raster)
require(maptools)
require(rgdal)
require(rasterVis)
require(rgeos)
require(polyclip)
a <- "MC"
    

```



```{r simulationAreas, echo=FALSE, cache=TRUE}
readURL <- "https://raw.githubusercontent.com/dcyr/LANDIS-II_IA_generalUseFiles/master/"
vegCodes <- read.csv(text = getURL(paste(readURL, "vegCodes.csv", sep="/")))
ecozones <- read.csv(text = getURL(paste(readURL, "ecoNames.csv", sep="/")))
## study area extent
readURL <- "https://github.com/dcyr/LANDIS-II_IA_generalUseFiles/raw/master/LandisInputs/"
tmpFile <- tempfile()
url <- paste(readURL, a, "/landtypes_", a, ".tif", sep="")
download.file(url, tmpFile, method="wget")
simArea <- raster(tmpFile)
landtypesN <- length(unique(values(simArea)[which(is.na(values(simArea)) == F &  values(simArea) != 0)]))
```



### Context

This document presents pixel-level simulations of one specific land type chosen among the `r landtypesN` land types defined for that quadrat. It was subjectively chosen among the most common land types to represent one general type of ecosystem.

We use such pixel-level simulation to assess the realism of emerging succession as simulated by _LANDIS-II Biomass Succession_, to minimise biases and document remaining ones.


```{r ecozones, echo=FALSE, cache=TRUE}
### fetching ecozones shapefile online
url <- "http://sis.agr.gc.ca/cansis/nsdb/ecostrat/zone/ecozone_shp.zip"
###
tmpFile <- tempfile()
tmpDir <- tempdir()
download.file(url, tmpFile)
### unzipping
fileNames <- unzip(tmpFile, list=TRUE)[,1]
unzip(zipfile = tmpFile,  files = fileNames, exdir=tmpDir)
### reading shapefile 
fileIndex <- grep("\\.", basename(fileNames))
zipFolder <- basename(fileNames[-fileIndex])
shapefileName <- basename(fileNames)[fileIndex]
shapefileName <- unique(as.character(lapply(strsplit(shapefileName, "\\."), function(x) x[[1]])))
### loading shapefile
shapeFile <- paste(tmpDir, zipFolder, shapefileName, sep="/")
ecozoneMap <- readShapeSpatial(shapeFile, proj4string = CRS("+proj=longlat +datum=WGS84"))
## renaming and reordering factors
levels(ecozoneMap$ZONE_NAME)[which(levels(ecozoneMap$ZONE_NAME)=="Boreal PLain")] <- "Boreal Plain"
ecozoneMap$ZONE_NAME <- factor(ecozoneMap$ZONE_NAME, levels=c("MixedWood Plain",
                                                              "Atlantic Maritime",
                                                              "Boreal Shield",
                                                              "Prairie",
                                                              "Boreal Plain",
                                                              "Montane Cordillera",
                                                              "Pacific Maritime",
                                                              "Boreal Cordillera",
                                                              "Taiga Shield",
                                                              "Taiga Plain",
                                                              "Taiga Cordillera",
                                                              "Hudson Plain",
                                                              "Southern Arctic",
                                                              "Northern Arctic",
                                                              "Arctic Cordillera"))

###  specifying colors
cols <- c("palegreen",
           "seagreen3",
           "palegreen4",
           "navajowhite1",
           "darkseagreen2",
           "darkkhaki",
           "darkolivegreen4",
           "burlywood4",
           "bisque3",
           "burlywood3",
           "darkslategray4",
           "lemonchiffon4",
           "lightblue1",
           "lightcyan1",
           "lightblue3")
names(cols) <- levels(ecozoneMap$ZONE_NAME)
```


```{r canadaMap, echo=FALSE, cache=TRUE}
can1 <- getData('GADM', country="CAN", level=1) ## level 1: provinces 
```

```{r miniGIS, echo=FALSE, cache=TRUE}
# reference projection
crsArea <- CRS("+init=epsg:3153") ### NAD83(CSRS) / BC AlbersAlbers Equal Area
# projecting maps
can1<- spTransform(can1, crsArea)# NAD83(CSRS) / Statistics Canada Lambert
ecozoneMap <- spTransform(ecozoneMap, crsArea)
simArea <- projectRaster(simArea, crs = crsArea)
# reference extent
bc <- can1[can1$NAME_1 == "British Columbia",]
e <-  extent(bc)
# adding 100 km around province extent
e <- e+100000
# cropping shapefiles
can1Cropped <- crop(can1, e, byid=TRUE)
# ### rasterize ecozones
canRaster <- ecoRaster <- raster(ncol=1000, nrow=1000)
extent(canRaster) <- extent(ecoRaster) <- extent(can1Cropped)
#
canRaster <- rasterize(can1Cropped, canRaster)
ecoRaster <- rasterize(ecozoneMap, ecoRaster, field="ZONE_NAME")
ecoRaster <- resample(ecoRaster, canRaster, method='ngb')
ecoRaster[is.na(canRaster)] <- NA
#

ecoRaster <- ratify(ecoRaster)
x <- levels(ecoRaster)[[1]]
x$ecoName <- levels(ecozoneMap$ZONE_NAME)[x$ID]
x$cols <- as.character(cols)[x$ID] 
#x$ID <- 1:nrow(x)
levels(ecoRaster) <- x
```

```{r mapPlot, echo=FALSE, cache=TRUE, fig.height=5, fig.width=7, fig.align="centre"}
### plot 
breakpoints <- unique(values(ecoRaster))
breakpoints <- breakpoints[is.na(breakpoints)==F]
breakpoints <- breakpoints[order(breakpoints)]
breakpoints <- c(min(breakpoints)-1, breakpoints)
#
### empty plot
r <- ecoRaster
r <- setExtent(r, ext=e, keepres=TRUE, snap=TRUE)
r[] <- NA
plot(r, box=FALSE, axes=F, frame.plot=F, xaxt='n', yaxt='n', main= paste0("Simulated quadrat - ", ecozones[which(ecozones$code == a), "name"]))
### ecozone plot
plot(ecoRaster, breaks=breakpoints,  col=x$cols, legend=FALSE, add=TRUE)
### Provinces boundaries
plot(can1Cropped, border="grey25", lwd=1, add=TRUE)
### simulation areas
plot(simArea, add=TRUE, col="grey25", legend=FALSE)

### LatLong gridlines
# llgridlines(can1Cropped, col="black", cex=0.6)
### legend
legend("topright", legend = rev(levels(ecoRaster)[[1]]$ecoName), fill = rev(x$col), cex=0.7, bty="n", title = "Ecozones")
```

-------

### Southern Interior Mountains - Interior Cedar Hemlock 
```{r URLs, echo = F, cache = F}
location <- "../"
results <- paste0(location, "Figures/multiSppLandisSite_MC_987_4.png")
```

For that type of ecosystem, the following species were included.

1. Western Hemlock (TSUG.HET)
2. Western Redcedar (THUJ.PLI)
3. Douglas-Fir (PSEU.MEN)
4. Trembling Aspen (POPU.TRE)
5. Spruces (PICE.ENG, but includes all types of spruces except Black Spruce)
6. Lodgepole Pine (PIN.CON.LAT)


* In the figure below, note that some panels present average values AND enveloppes, while others only present average values.

* The number of cohorts lumped into 10-yrs age classes is used as a proxy for **structural complexity**. That is only an indicator as some cohorts may be present in negligible amounts, especially those that established under already dense canopy.

* In _LANDIS-II Biomass Succession_ the percentage of maximum aboveground biomass occupied thoughout the simulation is used as a **proxy for shade**. Shade thresholds must be defined by the user to create shade classes that may limit the establishment of species based on their shade tolerance. The shade tolerance and other important traits of all species included in landscape-scale simulations are presented in the table below. As our simulations are currently parameterized, a given species establishment becomes limited when site shade classe reaches a value equal to shade tolerance of that species. Establishement becomes impossible when shade class exceed it.

-------

### Simulation setup

* A combination of 5 or 6 species start to growth from bare ground and interact for 1000 years with unlimited seed supply. The emerging succession is presented below.

* One cohort of each species is established at the beginning of each simulation. That is important to consider as that may differ from what mostly occur in nature, where cohort establishment may be delayed for some species, especially after large-scale disturbances and where species are distributed in highly contageous manners.

* Because seed sources are not limiting in those simulations, any locally extinct species can come back later in the simulations.

-------

### Pixel-level simulation results
![alt text](`r results`)


```{r speciesTraits, echo=FALSE}
readURL <- "https://raw.githubusercontent.com/dcyr/LANDIS-II_IA_generalUseFiles/master/LandisInputs/"
species <- read.table(text = getURL(paste0(readURL, a, "/species_", a, ".txt")),
                       skip=1, blank.lines.skip=T, comment.char=">")
rownames(species) <- species[,1]
species[,1] <- vegCodes[match(species[,1], vegCodes$LandisCode), "commonName"]
species <- species[, c(1,2,3,4)]

colnames(species) <- c("CommonName", "MaxLongevity", "SexualMaturity", "ShadeTolerance")
```

-------

### Simulated species

Only a subset of the following are included in the pixel-level simulations presented above. However, the following listed species are all those included in landscape-level simulations.

```{r, echo = FALSE}
species
```
